---
title: Intro to Histograms
author: Chris Makowski
date: '2017-10-02'
slug: intro-to-histograms
categories: []
tags:
  - Data Science
  - Histograms
  - Data
---
Let's construct a histogram! First, let's upload the three essential libraries: Lahman, sqldf, and ggplot2. Next, let's establish and run a query to see the difference in weight amongst those in the MLB. Once we have set up and ran our query we will find the different weights of all the players in a column in the console of Rstudio. However, wouldn't it be nice to see the results in a visual manner? How about a histogram?
```{r}
library(Lahman)
library(ggplot2) 
library(sqldf) 
# extracting the data-------------------

query<-"SELECT weight
FROM Master"

result<-sqldf(query)
```
In order to set up a histogram, we need to run a 'ggplot,' to do this, and allow for further modifications on the histogram, we must run the ggplot as 'ggplot()+.' On the next line, we run 'geom_histogram' and plug in the data of the histogram inside following parentheses such as the weight itself 'data=result' and their accompanying bars, the color of the bars that represent the data'color=purple,' 'fill=green,' the syntax of the x-axis 'aes(x=weight),' and the specificity of the weight on the x-axis 'bins=40.' Finally, to put your mark on the histogram, put a '+' sign at the end of the preceding line of syntax, to allow for a title to be placed on the histogram. On the next line run 'ggtitle' and in parentheses what you would like the title to be, in this case I used 'Body Weight Distribution for Baseball Players.'

```{r}
# visualizing the data----------------

ggplot()+
  geom_histogram(data=result,aes(x=weight),color="purple",fill="green",bins=40)+
  ggtitle("Body Weight Distribution for Baseball Players")

```

And that is how you do it! If you have any questions feel free to contact me and tune in for more exciting posts about the wonderous world of Rstudio and Data Science!
